---
title: "COVID-19 thrombosis analysis"
author: "Reed Sorensen"
date: "22 March 2022"
output: 
  html_document:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Tables by highest dosage of anti-coagulation therapy {.tabset}

```{r include=FALSE, message=FALSE}

library(dplyr)
library(lubridate)
library(arrow)
library(knitr)
library(kableExtra)
library(purrr)


code_dir <- "/mnt/workspace/GWTG/COVID19/ihme_code/thrombosis/aha_covid_anticoagulation/"
config2_path <- c(
  "/mnt/workspace/GWTG/COVID19/ihme_diagnostics/thrombosis/working_dirs/v2022-03-22/processed_data/config2.RDS"
)
config2 <- readRDS(config2_path)

output_dir <- file.path(config2$working_dir, "results")
if (!dir.exists(output_dir)) dir.create(output_dir)


df_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/vars.parquet"))
df_newvars <- read_parquet(file.path(config2$working_dir, "variable_specs/newvars.parquet"))
df_table1_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/table1_vars.parquet"))
df_table3_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/table3_vars.parquet"))

df1 <- read_parquet(file.path(config2$working_dir, "processed_data/df1.parquet"))
df2 <- read_parquet(file.path(config2$working_dir, "processed_data/df2.parquet"))
df4_tmp <- read_parquet(file.path(config2$working_dir, "processed_data/df4_tmp.parquet"))
df4 <- read_parquet(file.path(config2$working_dir, "processed_data/df4.parquet"))


source(file.path(code_dir, "99_analytic_functions.R"))


df_vars2 <- df_vars[, c("new_var", "type")] %>% 
  mutate(stage = "orig") %>%
  filter(!duplicated(.))

df_newvars2 <- df_newvars[, c("new_variable_name", "type")] %>%
  mutate(stage = "derived") %>%
  filter(!duplicated(.))
  
df_tmp <- df_table1_vars %>%
  left_join(df_vars2, by = c("variable_name" = "new_var")) %>%
  left_join(df_newvars2, by = c("variable_name" = "new_variable_name")) %>%
  mutate(type2 = ifelse(is.na(type.x), type.y, type.x))

options(scipen = 999)

```

#### Any


```{r echo=FALSE, message=FALSE}

anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general"
)

anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general",
  nonmissing_denom = FALSE
)


```


#### Before ICU

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)


anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)



  
```

#### Before health event

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe"
)


anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe", 
  nonmissing_denom = FALSE
)

```


#### Any (post July 2020) {.active}


```{r echo=FALSE, message=FALSE}

anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general"
)

anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general",
  nonmissing_denom = FALSE
)


```


#### Before ICU (post July 2020)

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)


anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)



  
```

#### Before health event (post July 2020)

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe"
)


anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe", 
  nonmissing_denom = FALSE
)

```




### Missingness

Table 3, missingness -- Variables with "NC"
```{r echo=FALSE, message=FALSE}


# df5 <- as.data.frame(cbind(
#   df1, 
#   data.frame(during_ami_any = df2[, "during_ami_any"], admission_date = df2[, "admission_date"]) )) %>%
#   mutate(pid = paste0(FACILITY_DISPLAY_ID, "__", PATIENT_DISPLAY_ID)) %>%
#   arrange(pid, admission_date)
# 
# # head(df5$pid)
# # head(df4_tmp$pid)
# # tail(df5$pid)
# # tail(df4_tmp$pid)
# df5$meets_exclusion_criteria <- df4_tmp$meets_exclusion_criteria
# df5$is_earliest_admission_date <- df4_tmp$is_earliest_admission_date
# df5$has_duplicate_adm_dates <- df4_tmp$has_duplicate_adm_dates
#     
# df6 <- df5 %>% 
#   filter(!has_duplicate_adm_dates) %>%
#   filter(is_earliest_admission_date) %>%
#   filter(!meets_exclusion_criteria %in% 1)

df6 <- df4

df_tbl3_1 <- df_table3_vars %>% filter(!is.na(nc_level))

dat3_1 <- bind_rows(lapply(1:nrow(df_tbl3_1), function(i) {
  if (FALSE) {
    i <- 1
  }
  df_i <- df_tbl3_1[i,]
  var_tmp <- df_i[1, "variable_name"]
  label_tmp <- df_i[1, "label"]
  
  df_yes_tmp <- df6[df6[, var_tmp] %in% 1, ]
  
  df_out <- data.frame(stringsAsFactors = FALSE,
    Variable = label_tmp,
    `Original variables` = paste0(var_tmp, ", ", df_i[1, "date_var"]),
    No = sum(df6[, var_tmp] %in% 2),
    NC = sum(df6[, var_tmp] %in% df_i[1, "nc_level"]),
    Missing = sum(is.na(df6[, var_tmp])),
    Yes = sum(df6[, var_tmp] %in% 1),
    `If yes, non-missing date` =  sum(!is.na(df_yes_tmp[, df_i[, "date_var"]]))
  )
  return(df_out)
}))


df_tbl3_2 <- df_table3_vars %>% filter(!is.na(nond_level))

dat3_2 <- bind_rows(lapply(1:nrow(df_tbl3_2), function(i) {
  if (FALSE) {
    i <- 7
  }
  df_i <- df_tbl3_2[i,]
  var_tmp <- df_i[1, "variable_name"]
  label_tmp <- df_i[1, "label"]
  cat(label_tmp, "\n")
  
  df_yes_tmp <- df6[df6[, var_tmp] %in% 1, ]
  
  df_out <- data.frame(stringsAsFactors = FALSE,
    Variable = label_tmp,
    `Original variables` = paste0(var_tmp, ", ", df_i[1, "date_var"]),
    `No/ND` = sum(df6[, var_tmp] %in% df_i[1, "nond_level"]),
    Missing = sum(is.na(df6[, var_tmp])),
    Yes = sum(df6[, var_tmp] %in% 1) )
  
  # if (var_tmp == "CLNBLDTRANS") {
  if (0) {
    df_out[, "If yes, non-missing date"] <- NA
  } else {
    df_out[, "If yes, non-missing date"] =  sum(!is.na(df_yes_tmp[, df_i[, "date_var"]]))
    
  }
  return(df_out)
}))


kbl(dat3_1) %>%
  kable_classic()


```

Table 3, missingness -- Variables with "No/ND"

```{r echo=FALSE, message=FALSE}
kbl(dat3_2) %>%
  kable_classic()
```



### Extra stuff (14 March 2022)

#### 1. Among people with 'any_during_anticoag_whatsoever'==0, how much missingness among A/C variables

```{r echo=TRUE, message=FALSE}

# a)  Include only patients with "No" or "Yes" with non-missing dates for anticoagulation AND  
# - this would apply to any anticoagulation choice


date_vars <- c(
  "during_anticoag_dvt_date", 
  "during_subq_lmwh_low_date", 
  "during_subq_lmwh_intermediate_date",
  "during_subq_lmwh_full_date", 
  "during_subq_uh_date", 
  "during_subq_paruh_date", 
  "during_doac_date", 
  "during_warfarin_date", 
  "during_argatroban_date", 
  "during_bivalirudin_date"
)

df4$n_nonmissing_anticoag_dates <- apply(df4[, date_vars], 1, function(x) {
  sum(!is.na(x))
})

anticoag_vars <- gsub("_date", "", date_vars)

df4$n_nonmissing_anticoag_values <- apply(df4[, anticoag_vars], 1, function(x) {
  sum(!is.na(x))
})

# table(df4$n_nonmissing_anticoag_values)
cat("\n\nAmong people with no recorded A/C usage whatsoever, the number of non-missing A/C values (among 10 variables)\n")
with(df4 %>% filter(any_during_anticoag_whatsoever == 0), table(n_nonmissing_anticoag_values))


df4$n_recorded_no <- apply(df4[, anticoag_vars], 1, function(x) {
  if (!all(is.na(x))) {
    sum(x == 0, na.rm = TRUE)
  } else {
    NA
  }
})

# table(df4$n_recorded_no, useNA = "always")
cat("\n\nAmong people with no recorded A/C usage whatsoever, the number of recorded 'No' observations (among 10 variables)\n")

with(df4 %>% filter(any_during_anticoag_whatsoever == 0), table(n_recorded_no, useNA = "always"))

# cat("\n\nAny A/C whatsoever")
# table(df4$any_during_anticoag_whatsoever, useNA = "always")
# 
# # length(date_vars)
# # with(df4, table(n_missing_anticoag_dates))
# 
# cat("\n\nAmong people taking A/C whatsoever, number of missing dates (10 dates total)\n")
# with(df4 %>% filter(any_during_anticoag_whatsoever == 1), table(n_missing_anticoag_dates))




```

#### 2. 

```{r echo=TRUE, message=FALSE}

# a)  Include only patients with "No" or "Yes" with non-missing dates for anticoagulation AND  
# - this would apply to any anticoagulation choice


date_vars <- c(
  "during_anticoag_dvt_date", 
  "during_subq_lmwh_low_date", 
  "during_subq_lmwh_intermediate_date",
  "during_subq_lmwh_full_date", 
  "during_subq_uh_date", 
  "during_subq_paruh_date", 
  "during_doac_date", 
  "during_warfarin_date", 
  "during_argatroban_date", 
  "during_bivalirudin_date"
)

df4$n_missing_anticoag_dates <- apply(df4[, date_vars], 1, function(x) {
  sum(is.na(x))
})

cat("\n\nAny A/C whatsoever")
table(df4$any_during_anticoag_whatsoever)

# length(date_vars)
# with(df4, table(n_missing_anticoag_dates))

cat("\n\nAmong people taking any A/C whatsoever, number of missing dates (10 dates total)\n")
with(df4 %>% filter(any_during_anticoag_whatsoever == 1), table(n_missing_anticoag_dates))


# how many people are on low dose DOAC
table(df4$during_anticoag_dvt_type)

df_tmp <- df4 %>%
  filter(during_anticoag_dvt_type == "Low Dose DOAC") %>%
  .[, c("ADMDT", "during_anticoag_dvt_date", anticoag_vars)]


```






### All variables before/after applying exclusion criteria

```{r echo=FALSE, message=FALSE}

iter_vars <- unique(df_vars$new_var)
iter_vars2 <- iter_vars[!iter_vars %in% ""]

# for (var_tmp in iter_vars2) {
for (var_tmp in c(iter_vars2, unique(df_newvars$new_variable_name))) {
  if (FALSE) {
    # var_tmp <- "during_doac"
    var_tmp <- "existing_afib"
  }
  
  if (var_tmp %in% iter_vars2) {
    try({
      
      df_tmp <- filter(df_vars, new_var == var_tmp)
      
      orig_tmp <- df_tmp[1, "orig_var"]
      # cat("\n\n\n\n", var_tmp, "\n\n")
      
      if (df_tmp[1, "type"] %in% c("Binary", "Categorical", "Unit label")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        print(table(df1[, orig_tmp], useNA = "always"))
        
        cat("\n\nAnalytic variable, pre-exclusion")
        print(table(df4_tmp[, var_tmp], useNA = "always"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        print(table(df4[, var_tmp], useNA = "always"))
      }
      
      if (df_tmp[1, "type"] %in% c("Continuous", "Integer")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        hist(df1[, orig_tmp], breaks = 30, main = df_tmp[1, "orig_var"])
        
        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
      if (df_tmp[1, "type"] %in% c("Date")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        hist(df1[, orig_tmp], breaks = 30, main = df_tmp[1, "orig_var"])
        
        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
    })
    
  } else {
      
    try({
      
      df_tmp <- filter(df_newvars, new_variable_name == var_tmp)

      cat("\n\n\n\n", var_tmp, "\n\n")

      if (df_tmp[1, "type"] %in% c("Binary", "Categorical", "Unit label")) {
        cat("\n\nAnalytic variable, pre-exclusion")
        print(table(df4_tmp[, var_tmp], useNA = "always"))

        cat("\n\nAnalytic variable, post-exclusion")
        print(table(df4[, var_tmp], useNA = "always"))
      }

      if (df_tmp[1, "type"] %in% c("Continuous", "Integer")) {

        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))

        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }

      if (df_tmp[1, "type"] %in% c("Date")) {

        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))

        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
    })
  }
  
}



```
