---
title: "COVID-19 thrombosis analysis"
author: "Reed Sorensen"
date: "23 May 2022"
output: 
  html_document:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Tables by highest dosage of anti-coagulation therapy {.tabset}
```{r include=FALSE, message=FALSE}

library(dplyr)
library(lubridate)
library(arrow)
library(knitr)
library(kableExtra)
library(purrr)


code_dir <- "/mnt/workspace/GWTG/COVID19/ihme_code/thrombosis/aha_covid_anticoagulation/"
config2_path <- c(
  "/mnt/workspace/GWTG/COVID19/ihme_diagnostics/thrombosis/working_dirs/v2022-05-23__5/processed_data/config2.RDS"
)
config2 <- readRDS(config2_path)

output_dir <- file.path(config2$working_dir, "results")
if (!dir.exists(output_dir)) dir.create(output_dir)


df_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/vars.parquet"))
df_newvars <- read_parquet(file.path(config2$working_dir, "variable_specs/newvars.parquet"))
df_table1_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/table1_vars.parquet"))
df_table3_vars <- read_parquet(file.path(config2$working_dir, "variable_specs/table3_vars.parquet"))

df1 <- read_parquet(file.path(config2$working_dir, "processed_data/df1.parquet"))
df2 <- read_parquet(file.path(config2$working_dir, "processed_data/df2.parquet"))
df4_tmp <- read_parquet(file.path(config2$working_dir, "processed_data/df4_tmp.parquet"))
df4 <- read_parquet(file.path(config2$working_dir, "processed_data/df4.parquet"))


source(file.path(code_dir, "99_analytic_functions.R"))


df_vars2 <- df_vars[, c("new_var", "type")] %>% 
  mutate(stage = "orig") %>%
  filter(!duplicated(.))

df_newvars2 <- df_newvars[, c("new_variable_name", "type")] %>%
  mutate(stage = "derived") %>%
  filter(!duplicated(.))
  
df_tmp <- df_table1_vars %>%
  left_join(df_vars2, by = c("variable_name" = "new_var")) %>%
  left_join(df_newvars2, by = c("variable_name" = "new_variable_name")) %>%
  mutate(type2 = ifelse(is.na(type.x), type.y, type.x))

options(scipen = 999)

```

```{r include=FALSE, message=FALSE}

# # check on how AFIBPRES is processed for the analytic dataset
# # appears to have few 'yes' values in the table below:
# #   Table 3, missingness -- Variables with "No/ND"
# 
# df_newvars2 <- df_newvars %>%
#   filter(!new_variable_name %in% "")
# exvar_idx <- which(df_newvars2$new_variable_name == "meets_exclusion_criteria")
# exclusion_vars <- df_newvars2$new_variable_name[1:exvar_idx]
# 
# table(df1$AFIBPRES)
# table(df2$AFIBPRES)
# table(df4_tmp$AFIBPRES)
# table(df4$AFIBPRES)
# 
# keep_vars_tmp <- c("AFIBPRES", "AFIBPRESDT", "during_afib", "during_afib_date", "ADMDT", exclusion_vars)
# df_tmp1 <- df4_tmp %>%
#   filter(AFIBPRES == 1) %>%
#   as.data.frame(.) %>%
#   .[, keep_vars_tmp]
# 
# 
# df_tmp2 <- df4_tmp %>%
#   filter(during_afib == 1) %>%
#   select(during_afib, during_afib_date, admission_date, has_afib_at_admission) %>%
#   mutate(
#     tmp0 = during_afib %in% 1 & during_afib_date %in% c(NA, admission_date),
#     tmp1 = during_afib %in% 1,
#     tmp2 = during_afib_date %in% c(NA, admission_date),
#     tmp2a = during_afib_date %in% c(NA),
#     tmp2b = during_afib_date %in% c(admission_date),
#     tmp2c = during_afib_date == c(admission_date),
#     tmp3 = during_afib %in% 1 & (during_afib_date == admission_date | is.na(during_afib_date)),
#   )
# 
# df_tmp3 <- df_tmp2 %>%
#   filter(is.na(during_afib_date))

```



#### Any


```{r echo=FALSE, message=FALSE}

anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general"
)

anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general",
  nonmissing_denom = FALSE
)


```


#### Before ICU

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)


anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)



  
```

#### Before health event

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe"
)


anticoag_table_binary(
  dat = df4,
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe", 
  nonmissing_denom = FALSE
)

```


#### Any (post July 2020) {.active}


```{r echo=FALSE, message=FALSE}

anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general"
)

anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_general",
  nonmissing_denom = FALSE
)


```


#### Before ICU (post July 2020)

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)


anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforeicu"
)



  
```

#### Before health event (post July 2020)

```{r echo=FALSE, message=FALSE}
anticoag_table_continuous(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe"
)


anticoag_table_binary(
  dat = df4 %>% filter(admission_date >= as_date("2020-08-01")),
  df_labels = df_tmp,
  type_variable = "type2",
  group_var = "highest_dose_beforehe", 
  nonmissing_denom = FALSE
)

```




### Missingness {.tabset}


```{r echo=FALSE, message=FALSE}


# df5 <- as.data.frame(cbind(
#   df1, 
#   data.frame(during_ami_any = df2[, "during_ami_any"], admission_date = df2[, "admission_date"]) )) %>%
#   mutate(pid = paste0(FACILITY_DISPLAY_ID, "__", PATIENT_DISPLAY_ID)) %>%
#   arrange(pid, admission_date)
# 
# # head(df5$pid)
# # head(df4_tmp$pid)
# # tail(df5$pid)
# # tail(df4_tmp$pid)
# df5$meets_exclusion_criteria <- df4_tmp$meets_exclusion_criteria
# df5$is_earliest_admission_date <- df4_tmp$is_earliest_admission_date
# df5$has_duplicate_adm_dates <- df4_tmp$has_duplicate_adm_dates
#     
# df6 <- df5 %>% 
#   filter(!has_duplicate_adm_dates) %>%
#   filter(is_earliest_admission_date) %>%
#   filter(!meets_exclusion_criteria %in% 1)


make_missingness_table <- function(df6) {
  
  if (0) {
    df6 <- df4
  }

  df_tbl3_1 <- df_table3_vars %>% filter(!is.na(nc_level))
  
  dat3_1 <- bind_rows(lapply(1:nrow(df_tbl3_1), function(i) {
    if (FALSE) {
      i <- 1
    }
    df_i <- df_tbl3_1[i,]
    var_tmp <- df_i[1, "variable_name"]
    label_tmp <- df_i[1, "label"]
    
    df_yes_tmp <- df6[df6[, var_tmp] %in% 1, ]
    
    df_out <- data.frame(stringsAsFactors = FALSE,
      Variable = label_tmp,
      `Original variables` = paste0(var_tmp, ", ", df_i[1, "date_var"]),
      No = sum(df6[, var_tmp] %in% 2),
      NC = sum(df6[, var_tmp] %in% df_i[1, "nc_level"]),
      Missing = sum(is.na(df6[, var_tmp])),
      Yes = sum(df6[, var_tmp] %in% 1),
      `If yes, non-missing date` =  sum(!is.na(df_yes_tmp[, df_i[, "date_var"]]))
    )
    return(df_out)
  }))
  
  
  df_tbl3_2 <- df_table3_vars %>% filter(!is.na(nond_level))
  
  dat3_2 <- bind_rows(lapply(1:nrow(df_tbl3_2), function(i) {
    if (FALSE) {
      i <- 1
    }
    df_i <- df_tbl3_2[i,]
    var_tmp <- df_i[1, "variable_name"]
    label_tmp <- df_i[1, "label"]
    # cat(label_tmp, "\n")
    
    df_yes_tmp <- df6[df6[, var_tmp] %in% 1, ]
    
    df_out <- data.frame(stringsAsFactors = FALSE,
      Variable = label_tmp,
      `Original variables` = paste0(var_tmp, ", ", df_i[1, "date_var"]),
      `No/ND` = sum(df6[, var_tmp] %in% df_i[1, "nond_level"]),
      Missing = sum(is.na(df6[, var_tmp])),
      Yes = sum(df6[, var_tmp] %in% 1) )
    
    # if (var_tmp == "CLNBLDTRANS") {
    if (0) {
      df_out[, "If yes, non-missing date"] <- NA
    } else {
      df_out[, "If yes, non-missing date"] =  sum(!is.na(df_yes_tmp[, df_i[, "date_var"]]))
      
    }
    return(df_out)
  }))
  
  
  out <- list(dat3_1 = dat3_1, dat3_2 = dat3_2)
  return(out)
  
}

```



#### Missingness -- restricted to post July 2020

```{r echo=FALSE, message=FALSE}
miss1 <- make_missingness_table(df6 = df4 %>% filter(admission_date >= as_date("2020-08-01")))

cat("\nMissingness -- Variables with 'NC'\n\n")
kbl(miss1$dat3_1) %>%
  kable_classic()


cat("\nMissingness -- Variables with 'No/ND'\n\n")
kbl(miss1$dat3_2) %>%
  kable_classic()


```

#### Missingness -- no time restriction {.active}

```{r echo=FALSE, message=FALSE}
miss2 <- make_missingness_table(df6 = df4)

cat("\nMissingness -- Variables with 'NC'\n\n")
kbl(miss2$dat3_1) %>%
  kable_classic()


cat("\nMissingness -- Variables with 'No/ND'\n\n")
kbl(miss2$dat3_2) %>%
  kable_classic()

```

### Extra stuff (14 March 2022)

#### 1. Among people with 'any_during_anticoag_whatsoever'==0, how much missingness among A/C variables

```{r echo=TRUE, message=FALSE}

# a)  Include only patients with "No" or "Yes" with non-missing dates for anticoagulation AND  
# - this would apply to any anticoagulation choice


date_vars <- c(
  # "during_anticoag_dvt_date", # removing prophylactic A/C for DVT; variable used intermittently
  "during_subq_lmwh_low_date", 
  "during_subq_lmwh_intermediate_date",
  "during_subq_lmwh_full_date", 
  "during_subq_uh_date", 
  "during_subq_paruh_date", 
  "during_doac_date", 
  "during_warfarin_date", 
  "during_argatroban_date", 
  "during_bivalirudin_date"
)

df4$n_nonmissing_anticoag_dates <- apply(df4[, date_vars], 1, function(x) {
  sum(!is.na(x))
})

anticoag_vars <- gsub("_date", "", date_vars)

df4$n_nonmissing_anticoag_values <- apply(df4[, anticoag_vars], 1, function(x) {
  sum(!is.na(x))
})
df4$n_missing_anticoag_values <- apply(df4[, anticoag_vars], 1, function(x) {
  sum(is.na(x))
})

# table(df4$n_nonmissing_anticoag_values)
cat("\n\nAmong people with no recorded A/C usage whatsoever, the number of non-missing A/C values (among 9 variables)\n")
with(df4 %>% filter(any_during_anticoag_whatsoever == 0), table(n_nonmissing_anticoag_values))


df4$n_recorded_no <- apply(df4[, anticoag_vars], 1, function(x) {
  if (!all(is.na(x))) {
    sum(x == 0, na.rm = TRUE)
  } else {
    NA
  }
})

# table(df4$n_recorded_no, useNA = "always")
cat("\n\nAmong people with no recorded A/C usage whatsoever, the number of recorded 'No' observations (among 9 variables)\n")

with(df4 %>% filter(any_during_anticoag_whatsoever == 0), table(n_recorded_no, useNA = "always"))

# cat("\n\nAny A/C whatsoever")
# table(df4$any_during_anticoag_whatsoever, useNA = "always")
# 
# # length(date_vars)
# # with(df4, table(n_missing_anticoag_dates))
# 
# cat("\n\nAmong people taking A/C whatsoever, number of missing dates (10 dates total)\n")
# with(df4 %>% filter(any_during_anticoag_whatsoever == 1), table(n_missing_anticoag_dates))


cat("\n\nAmong people with at least 1 recorded A/C, the number of non-missing A/C values (among 9 variables)\n")
# NOTE: this table includes several '0' non-missing A/C values
# this is because more A/C variables are used in the construction of 'any_during_anticoag_whatsoever' variable than
# those listed above. the ones listed above are only those with corresponding dates. please see lines 123 to 128 here:
# https://github.com/rsoren/aha_covid_dataprep_specs/blob/main/02_create_new_variables.csv
with(df4 %>% filter(any_during_anticoag_whatsoever == 1), table(n_nonmissing_anticoag_values))


```


#### 2. Other missingness questions

```{r}
cat("\n\nFor those that had any missing anticoagulation (OR yes but a missing date), how many of the anticoagulation fields were missing?\n\n")



for (var in anticoag_vars) {
  newvar_name <- paste0(var, "__yes_butnodate")
  df4[newvar_name] <-  df4[, var] %in% 1 & is.na(df4[, paste0(var, "_date")])
}
newvar_names <- names(df4)[grepl("__yes_butnodate", names(df4))]

sapply(newvar_names, function(x) table(df4[, x]), simplify = FALSE)

df4$n_yes_anticoag_but_missing_date <- apply(X = df4[, newvar_names], MARGIN = 1, FUN = function(x) {
  sum(x, na.rm = TRUE)
})

table(df4$n_missing_anticoag_values)
table(df4$n_yes_anticoag_but_missing_date)
table(df4$n_yes_anticoag_but_missing_date, df4$n_missing_anticoag_values)

# include only those with at least 1 missing anticoagulation OR at least 1 reported-anticoagulation-but-no-date
dfe1 <- df4 %>%
  filter(n_missing_anticoag_values > 0 | n_yes_anticoag_but_missing_date > 0) 

table(dfe1$n_missing_anticoag_values)



cat("For those that had any missing health events (OR yes but a missing date), how many health event fields were missing?")

for (include_limb_ischemia in c(TRUE, FALSE)) {
  
  if (include_limb_ischemia) {
    exclude_vars <- c("PATMANICU")
    cat("\n\n\nINCLUDING LIMB ISCHEMIA\n")
  } else {
    exclude_vars <- c("PATMANICU", "ACUTLMBISCH")
    cat("\n\n\nEXCLUDING LIMB ISCHEMIA\n")
  }
  
  df_healthvars <- df_table3_vars %>%
    filter(is.na(nc_level)) %>% # filtering on missing NC level because those happen to correspond to health events (except the ICU variable)
    filter(!variable_name %in% exclude_vars)
  
  healthvar_names <- character(nrow(df_healthvars))
  for (i in 1:nrow(df_healthvars)) {
    if (0) {
      i <- 1
    }
    var_tmp <- df_healthvars[i, "variable_name"]
    datevar_tmp <- df_healthvars[i, "date_var"]
    newvar_name <- paste0(var_tmp, "__yes_butnodate")
    df4[newvar_name] <-  df4[, var_tmp] %in% 1 & is.na(df4[, datevar_tmp])
    healthvar_names[i] <- newvar_name
  }
  
  df4$n_yes_healthevent_but_missing_date <- apply(X = df4[, healthvar_names], MARGIN = 1, FUN = function(x) {
    sum(x, na.rm = TRUE)
  })
  
  df4$n_missing_healthevents <- apply(df4[, df_healthvars$variable_name], 1, function(x) {
    sum(is.na(x))
  })
  
  cat("\n\nloop through all health events, summary stats\n")
  print(sapply(df_healthvars$variable_name, function(x) table(df4[, x], useNA = "always"), simplify = FALSE))
  
  cat("\n\nTables of missing health events; yes but missing data; and a cross-table of the two\n")
  print(table(df4$n_missing_healthevents))
  print(table(df4$n_yes_healthevent_but_missing_date))
  print(table(df4$n_yes_healthevent_but_missing_date, df4$n_missing_healthevents))
  
  dfe2 <- df4 %>%
    filter(n_missing_healthevents > 0 | n_yes_healthevent_but_missing_date > 0)
  
  cat("\n\nNumber of missing health events, among those with at least 1 missing OR yes-but-missing-date\n")
  print(table(dfe2$n_missing_healthevents, useNA = "always"))
  
}
# May need to look at this with and without acute limb ischemia as a health event to get a sense of the degree of overlap


```

#### 3. Misc.

```{r echo=TRUE, message=FALSE}


# how many people are on low dose DOAC
table(df4$during_anticoag_dvt_type)

# dfe <- df4 %>%
#   filter(during_anticoag_dvt_type == "Low Dose DOAC") %>%
#   .[, c("ADMDT", date_vars)]


```






### All variables before/after applying exclusion criteria

```{r echo=FALSE, message=FALSE}

iter_vars <- unique(df_vars$new_var)
iter_vars2 <- iter_vars[!iter_vars %in% ""]

# for (var_tmp in iter_vars2) {
for (var_tmp in c(iter_vars2, unique(df_newvars$new_variable_name))) {
  if (FALSE) {
    # var_tmp <- "during_doac"
    var_tmp <- "existing_afib"
  }
  
  if (var_tmp %in% iter_vars2) {
    try({
      
      df_tmp <- filter(df_vars, new_var == var_tmp)
      
      orig_tmp <- df_tmp[1, "orig_var"]
      # cat("\n\n\n\n", var_tmp, "\n\n")
      
      if (df_tmp[1, "type"] %in% c("Binary", "Categorical", "Unit label")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        print(table(df1[, orig_tmp], useNA = "always"))
        
        cat("\n\nAnalytic variable, pre-exclusion")
        print(table(df4_tmp[, var_tmp], useNA = "always"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        print(table(df4[, var_tmp], useNA = "always"))
      }
      
      if (df_tmp[1, "type"] %in% c("Continuous", "Integer")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        hist(df1[, orig_tmp], breaks = 30, main = df_tmp[1, "orig_var"])
        
        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
      if (df_tmp[1, "type"] %in% c("Date")) {
        cat("\n\nOriginal variable, pre-exclusion:", df_tmp[1, "orig_var"])
        hist(df1[, orig_tmp], breaks = 30, main = df_tmp[1, "orig_var"])
        
        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))
        
        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
    })
    
  } else {
      
    try({
      
      df_tmp <- filter(df_newvars, new_variable_name == var_tmp)

      cat("\n\n\n\n", var_tmp, "\n\n")

      if (df_tmp[1, "type"] %in% c("Binary", "Categorical", "Unit label")) {
        cat("\n\nAnalytic variable, pre-exclusion")
        print(table(df4_tmp[, var_tmp], useNA = "always"))

        cat("\n\nAnalytic variable, post-exclusion")
        print(table(df4[, var_tmp], useNA = "always"))
      }

      if (df_tmp[1, "type"] %in% c("Continuous", "Integer")) {

        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))

        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }

      if (df_tmp[1, "type"] %in% c("Date")) {

        cat("\n\nAnalytic variable, pre-exclusion")
        hist(df4_tmp[, var_tmp], breaks = 30, main = paste0(var_tmp, ", pre-exclusion"))

        cat("\n\nAnalytic variable, post-exclusion")
        hist(df4[, var_tmp], breaks = 30, main = paste0(var_tmp, ", post-exclusion"))
      }
      
    })
  }
  
}



```
